Inferring probabilistic stellar rotation periods using Gaussian processes

### 2017-02-02 documents/ms.tex
removed references in abstract

\begin{abstract}
The light curves of spotted, rotating stars are often non-sinusoidal and
quasi-periodic---spots are not static on the stellar surface and have finite
lifetimes, causing stellar flux variations to slowly shift in phase.
A strictly periodic sinusoid therefore cannot accurately model a 
rotationally modulated stellar light curve.
Physical models of the stellar surface also have many drawbacks preventing effective
inference, such as having highly degenerate or high-dimensional parameter spaces
\citep[\eg][]{Russell1906, Jeffers2009, Kipping2012}.
In this work, we introduce an appropriate {\it effective} model: a Gaussian Process (GP)
with a quasi-periodic covariance kernel function.
Using this highly flexible model, 
we are able to sample from the posterior Probability Density Function
(PDF) of the periodic parameter, marginalising over the other kernel hyperparameters
using a Markov Chain Monte Carlo approach.
To test this method's effectiveness, we infer rotation periods from 
\naigrain\ simulated light curves, demonstrating that our results 
    are more accurate than both a sine-fitting periodogram and an 
    AutoCorrelation Function (ACF) method.
We also demonstrate that it works effectively on real data, by 
inferring rotation periods for \nkois\ \Kepler\ stars.
Because this method delivers posteriors PDFs, 
it will enable other studies involving stellar rotation, 
such as inferring stellar
ages or stellar obliquities in exoplanet systems, or characterising star-planet interactions.
% Thehe ACF method is not applicable to unevenly spaced data,
% therefore it is inappropriate for the ~$10^7$ stellar light curves expected
% to be produced by \LSST\ unlike the GP method.
% Furthermore, the improvement is expected to be even more dramatic when applied
% to real, noisy {\it Kepler} light curves, since the GP method is well suited
% to modelling rotation signals and correlated noise simultaneously.

>>>>

\begin{abstract}
The light curves of spotted, rotating stars are often non-sinusoidal and
quasi-periodic---spots are not static on the stellar surface and have finite
lifetimes, causing stellar flux variations to slowly shift in phase.
A strictly periodic sinusoid therefore cannot accurately model a 
rotationally modulated stellar light curve.
Physical models of the stellar surface also have many drawbacks preventing effective
inference, such as having highly degenerate or high-dimensional parameter spaces.
% \citep[\eg][]{Russell1906, Jeffers2009, Kipping2012}.
In this work, we introduce an appropriate {\it effective} model: a Gaussian Process (GP)
with a quasi-periodic covariance kernel function.
Using this highly flexible model, 
we are able to sample from the posterior Probability Density Function
(PDF) of the periodic parameter, marginalising over the other kernel hyperparameters
using a Markov Chain Monte Carlo approach.
To test this method's effectiveness, we infer rotation periods from 
\naigrain\ simulated light curves, demonstrating that our results 
    are more accurate than both a sine-fitting periodogram and an 
    AutoCorrelation Function (ACF) method.
We also demonstrate that it works effectively on real data, by 
inferring rotation periods for \nkois\ \Kepler\ stars.
Because this method delivers posteriors PDFs, 
it will enable other studies involving stellar rotation, 
such as inferring stellar
ages or stellar obliquities in exoplanet systems, or characterising star-planet interactions.
% Thehe ACF method is not applicable to unevenly spaced data,
% therefore it is inappropriate for the ~$10^7$ stellar light curves expected
% to be produced by \LSST\ unlike the GP method.
% Furthermore, the improvement is expected to be even more dramatic when applied
% to real, noisy {\it Kepler} light curves, since the GP method is well suited
% to modelling rotation signals and correlated noise simultaneously.
